Inverse energy transfer in three-dimensional quantum vortex flows

### 2025-10-30 revision-2/referee-response.tex

\documentclass[a4paper,10pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{cite}
\usepackage{amsmath}
\usepackage{graphicx}% Include figure files
\usepackage{bm}% bold math
\usepackage{float}
\usepackage{xcolor}
\def\red#1{\textcolor{black}{#1}}
\def\blue#1{\textcolor{blue}{#1}}

\def\refcomment#1{\textbf{\blue{\emph{#1}}}\\}
\hbadness=99999  % or any number >=10000

\def\red#1{\textcolor{red}{#1}}
\def\magenta#1{\textcolor{magenta}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\NOTE}[1]{\textbf{\color{red}[#1]}}
\newcommand*{\CHANGE}[1]{{\color{magenta}#1}}
\newcommand*{\DEL}[1]{{\color{green}#1}}

@@

We thank the reviewers for taking the time to read over and give constructive feedback on our paper titled “\emph{Inverse energy transfer in three-dimensional quantum vortex flows}", submitted to PRL. Please see below the point-by-point response to the comments made by the reviewers.

\subsection*{Referee A}

\refcomment{For issue 1), authors
commented “We omitted the plot of the superfluid helicity from the
original manuscript,” but I did not understand where or which part was
omitted.}

Originally we wanted to include the plot of the superfluid helicity in the plot of the normal fluid helicity. We decided against this as in this paper we decide to only focus on the normal fluid aspect. \CHANGE{Nevertheless, we have included a sentence in the manuscript emphasizing that the total helicity of the flow, normal fluid plus superfluid, is nor conserved in the reconnection process.}\\

@@

\refcomment{For issue 2), the total sign of helicity for two
reconnections is neutral; one is negative and the other is positive.
Therefore, total symmetry breaking does not occur. This means that a
lot of reconnections do not cause total symmetry breaking or chiral
imbalance in quantum turbulence. As a result, the inverse cascade will
not happen.}

%We would like to highlight one subtlety of orthogonal reconnection simulations, which is that the two reconnections occur asynchronously. Given a large enough separation between vortices within the pair, as was this case for the configuration in our work, the large timescales required to observe a reconnection can lead to the pairs having very different reconnection times. Small deviations build up over time and the minimum distance $\delta$ between the vortices within each pair is not identical. As a result, the two reconnections do not occur at the same time. Pre-reconnection, it is true that the there is no global symmetry breaking, as the normal fluid helicity is not only negligible but is cancelled out globally. In fact, even at the reconnection of the first pair of vortices, the normal fluid helicity generated by the second pair does not change in light of the reconnection and remains negligible. This can be seen by computing the evolution of the normal fluid helicity in half-boxes containing exactly one pair of vortices. Therefore we do indeed have global symmetry breaking, even though we say that we are only considering one vortex pair. If we extended the plot of helicity in Fig.2a, one would see that the helicity injected is of the negative sign, with the global helicity returning back to 0. \\

\CHANGE{
To generalise our findings, we perform the study of the chiral imbalance generated by reconnections occurring in Hopf-links, constituted by two, linked, orthogonal vortex rings,
whose initial relative position is varied (see new figures in Supplemental material). This allows to observe reconnections taking place at different angles.
Also in this geometrical setting, we observe a strong helicity injection, an increase of the integral scale of the flow, negative spectral fluxes and mutual friction force 
chirality. We have included the data referring to Hopf-link reconnections in the manuscript and in the Supplemental material.
Hence, we can claim, with supporting evidence, that reconnections trigger inverse energy transfer. However, this transfer of energy towards large scales may lead
to an inverse cascade only if reconnections are not exactly symmetrycal, \textit{i.e.} (simplifying) if the number of reconnections of a given type is not exactly balanced by 
the same number of reconnections between vortices of opposite circulation. We suggest that this might be the case when symmetry is broken by external forcing, as in thermal counterflow.}

%further to make this point, in Fig.2 of the paper we include the reconnection data of Hopf-links that constitute a range of reconnection angles. The helicity evolution shows a strong injection of helicity and an increase in the integral length scale post reconnection. A similar analysis that we did on the helical decomposition also showed the same inverse energy transfer, but less stronger than for the orthogonal reconnection. We hope that this evidence is enough for the referee to see that we do indeed have symmetry breaking in the system. \\

%Furthermore, we would like to remind the referee that the effect that we are reporting o occurs locally, and our results do not attempt to suggest that this mechanism is responsible for inverse energy cascades in 3D quantum turbulence such as in counterflow channels. However, we would like to refer the referee to recent experimental results in counterflow channels such as in Ref~\cite{marakov2015,gao2017,mastracci2019,yui2020}. In these works, the experimental team often report the observation of large-scale normal fluid turbulence in thermal counterflow, which is still unexplained. In Ref.\cite{mastracci2019} the authors suggest that the large scale normal fluid turbulence may result from wake structures. Using data taken from a recent preprint \cite{galantucci2025} that three of us were part of, we calculated that the the flux of normal fluid due to wake structure is positive, implying an energy transfer to \emph{smaller} scales rather than larger scales, suggesting that wake structures are not likely to be the reason behind the large-scale normal fluid turbulence observed. \\

@@

\refcomment{Regarding issue 3), a new normalization does not improve
the order of the kinetic energy spectrum and kinetic energy flux,
although Fig. 1 with dimensions in the response letter indeed shows a
reasonable order.}

In the revised manuscript, we have adapted the figures in the paper in order to use a normalization that is temperature independent scaling. %, which was the reason why the figures before showed an apparent inconsistency between the energy spectrum and the flux. 
The figures now show the same order as one would find when using dimensional units.

\refcomment{The finding of inverse energy transfer by single reconnection is
interesting. However, other events cause forward energy transfer of
normal fluid, e.g., nonlinear interaction, wake propagation, energy
dissipations by mutual friction and viscosity, and so on. Therefore,
it is unclear whether the inverse energy cascade of normal fluid is
generated by reconnections in quantum turbulence. This stage is
premature to recommend to Physical Review Letters.}

\CHANGE{We think that by including the data regarding Hopf-link reconnections, we generalised our findings beyond a particular type of reconnection. This inverse energy transfer,
triggered by classical-like phenomena (chiral imbalance, which as in classical turbulence produces energy flowing towards large scales), is strikingly dissimilar from 
the direct energy transfer observed in vortex reconnections occurring in classical viscous fluids. Our work shows that vortex reconnections in quantum fluids may possess 
different characteristics from the ones taking place in classical fluids, depending on the fluid component we focus on: if we analyse superfluid vortices, the approach and 
separation is akin to classical dynamics (see our recent work, Stasiak et al., PNAS Vol. 122 No. 21 e2426064122 (2025)); if we observe the normal fluid we observe a non-classical 
energy transfer. Furthermore, this work adds additional elements to the lively discussion in the community regarding similarities and differences between classical and quantum 
turbulence. We think that together with the suggestion that reconnections may trigger an inverse energy cascade, these are the main findings of our investigation, which are of 
relevance for a broad audience such as PRL's readers. We have slightly changed the conclusions to emphasise these aspects.}

\subsection*{Referee B}

%\begin{thebibliography}{4}
%    \bibitem{marakov2015}
%    A. Marakov, J. Gao, W. Guo, S.W. Van Sciver, G.G. Ihas, D.N. McKinsey, and W.F. Vinen, Visualization of the normal-fluid turbulence in counterflowing superfluid \textsuperscript{4}He, \emph{Phys. Rev. B} \textbf{91}, 094503 (2015).

%    \bibitem{gao2017}
%    J. Gao, E. Varga, W. Guo, and W.F. Vinen, Energy spectrum of thermal counterflow turbulence in superfluid helium-4, \emph{Phys. Rev. B} \textbf{96}, 094511 (2017)

%    \bibitem{mastracci2019}
%    B. Mastracci, S. Bao, W. Guo, and W.F. Vinen, Particle tracking velocimetry applied to thermal counterflow in superfluid \textsuperscript{4}He: motion of the normal fluid at small heat fluxes, \emph{Phys. Rev. Fluids} \textbf{4}, 083305 (2019).

%    \bibitem{yui2020}
%    S. Yui, H. Kobayashi, M. Tsubota, and W. Guo, Fully coupled dynamics of the two fluids in superfluid \textsuperscript{4}He: Anomalous anisotropic velocity fluctuations in counterflow, \emph{Phys. Rev. Lett.} \textbf{124}, 155301 (2020).

%    \bibitem{galantucci2025}
%    L. Galantucci, G. Krstulovic, and C. F. Barenghi, Quantum vortices leave a macroscopic signature in the normal fluid. \emph{arXiv preprint arXiv:2501.08309} (2025).

%\end{thebibliography}

### 2025-10-30 main_REV.tex
Major revisions to text- LG

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\NOTE}[1]{\textbf{\color{red}[#1]}}
\newcommand*{\CHANGE}[1]{{\color{brown}#1}}
\newcommand*{\DEL}[1]{{\color{green}#1}}

@@

\begin{figure*}
	\centering
  \includegraphics[width=\textwidth]{snapshots.pdf}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-4.pdf}

	% \end{subfigure} \hfill
  %   %
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	% \end{subfigure} \hfill
	\caption{
Three-dimensional rendering of the time evolution of
an initially orthogonal vortex configuration 
undergoing a vortex reconnection at $T = 1.9$K at dimensionless
times (from left to right)
$(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$, where $\tau_R = L_R/E^{1/2}_R$. The quantities $E_R$ and $L_R$ are the energy $E$ and integral length scales $L$ of the normal fluid at reconnection time $t_R$. The green
tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R=8.79\times10^{-4}$ . In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
    \label{fig:visualisation}
\end{figure*}

@@

In this work, we unveil a similar dynamics occurring in superfluid helium
($^4$He) as a result of vortex reconnections. 
%\red{The onset and development of quantum turbulence is fundamentally driven by such reconnections, which facilitates energy transfer and the breakdown of coherent vortex structures.}  
%Reconnections occur continuously in turbulence: they take place when
%two vortex lines collide and recombine, exchanging heads and tails, 
%altering the overall topology of the flow
%\cite{koplik-levine-1993,bewley-etal-2008,rorai-etal-2016,serafini-etal-2017,galantucci-baggaley-parker-barenghi-2019,villoisUniversalNonuniversalAspects2017,villois2020irreversible}. 
We show that the mutual friction force arising from the  
vortex reconnection is chiral, injecting in the normal fluid prevalently 
helicity of a given sign 
%\red{which is not globally conserved across the reconnection event}. 
Thus, as a consequence of vortex reconnections,
we observe an increase of the chiral imbalance of the quantum fluid, producing a transfer of kinetic energy from small to large scales, similarly to the phenomenology observed in 3D helical-decimated classical flows. 
 Unlike classical fluids, such a chiral imbalance arises naturally as physical process in the normal fluid.% because of vortex reconnections. 
%This is contrast to classical fluid dynamics,  where the helical characteristics of the flows producing 
% an inverse energy transfer require a careful synthetic construction 
% \cite{biferaleInverseEnergyCascade2012a,biferale-etal-2013,sahoo-etal-2017,plunianInverseCascadeEnergy2020a}.

@@

To model superfluid helium dynamics, we employ the recently developed FOUCAULT model
\cite{galantucciNewSelfconsistentApproach2020b}.  In this approach, superfluid vortex lines are
parametrized as one-dimensional space curves  $\s(\xi,t)$, $\xi$ and $t$ being arclength and time respectively, exploiting the large separation of length scales between the vortex core radius, the Lagrangian discretisation along the vortex lines $\Delta\xi$, and the average radius of curvature $R_c$ of the vortex lines. The vortex lines evolve according to the following equation of motion:
\begin{equation}
    \dot{\s}(\xi,t) = \v_s + \frac{\beta}{1+\beta}\left[\v_{ns}\cdot\s'\right]\s' + \beta\s'\times\v_{ns} + \beta'\s'\times\left[\s'\times\v_{ns}\right],
\end{equation}
where $\dot{\s} = \partial\s/\partial t$, $\s' = \partial\s/\partial\xi$ is the unit tangent vector, $\v_n$ and $\v_s$ are the normal fluid and superfluid velocities at $\s$, $\v_{ns} = \v_n-\v_s$, and $\beta,\, \beta'$ are temperature and Reynolds number dependent mutual friction coefficients \cite{galantucciNewSelfconsistentApproach2020b}. The calculation of the superfluid velocity $\v_s$ is performed via the computation of the Biot-Savart integral de-singularised with standard techniques (see Supplementary Material \cite{suppMat}). The normal fluid is described classically using the incompressible ($\nabla\cdot\v_n=0$) 
Navier-Stokes equation
\begin{equation}
    \frac{\partial\v_n}{\partial t} + (\v_n\cdot\nabla)\v_n = -\frac{1}{\rho}\nabla p  + \nu_n\nabla^2\v_n + \frac{\mathbf{F}_{ns}}{\rho_n} \; \; , 
\end{equation}
%
where $\rho_n$ and $\rho_s$ are the normal fluid and superfluid densities,
$\rho=\rho_n + \rho_s$,  $p$ is the pressure,  $\nu_n$ is the kinematic 
viscosity of the normal fluid, and the mutual friction force per unit
volume, $\mathbf{F}_{ns}$, is the line integral of the mutual friction 
force per unit length, $\mathbf{f}_{ns}$ \cite{suppMat}:
%
\begin{equation}
\mathbf{F}_{ns}(\x) = 
\oint_{\mathcal{C}}\delta(\x-\s)\mathbf{f}_{ns}(\s)d\xi,     
\end{equation}
$\mathcal{C}$ representing the entire vortex configuration. 
The regularisation of mutual friction is performed using a physically self-consistent scheme \cite{galantucciNewSelfconsistentApproach2020b}. %, gualtieri2015exact, gualtieri2017turbulence}. 
We consider a periodical box of size $2\pi$ (so that  wavevectors are integers).

To study the reconnection dynamics, we consider two pairs of initially orthogonal vortices (where the corresponding vortices of each pair have opposite circulation in order to preserve periodicity along the boundaries)
at two distinct temperatures, $T=1.9K$ and $T=2.1K$. The vortex pairs are separated by the distance $D_{\ell}$; each vortex within each pair is initially at distance $d_{\ell}$ to the other vortex, such that $d_{\ell}\ll D_{\ell}$ to ensures that the dynamics in the vicinity of the reconnection is dominated by local interactions, and that the far-field contribution from the other vortex pair is negligible. 
All results reported hereafter refer only to the reconnection of one pair of vortices. 
%\red{In our simulations, the reconnections of the two pairs of vortices occur at different times; here we consider only the vortex dynamics up to and shortly after the first reconnection (see Fig.~\ref{fig:total-helicity}). Hence, all results reported and discussed herafter refer to the first reconnection.}.

The evolution of the vortex reconnection of a single pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by mutual friction, here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2025experimental}. The second row shows the rendering of the local helicity $H(\x)=\v_n\cdot\bom$, where we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during reconnection there is a net 
sudden normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity}. We will come back to this finding later.
\begin{figure}[t]
    \centering

      \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{helicity_total.pdf}
        \caption{}\label{fig:}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{integral-scale-scaled.pdf}
        \caption{}\label{fig:}

    \end{subfigure}
\caption{\blue{Temporal evolution of the normal fluid helicity $\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$. For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$.}
%\red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the two reconnections occurring at different times; same colour scheme as in Fig.~\ref{fig:visualisation}.}
}
\label{fig:total-helicity}
\end{figure}

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection. The energy spectrum $E(k)$ is displayed in Fig.~3a.
\begin{figure}
    \centering
    % \includegraphics[width=0.49\textwidth]{energy-flux-spec.pdf}

    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{energy-spec-scaled.pdf}
        \caption{}\label{fig:kinetic-energy}
    \end{subfigure}
    \begin{subfigure}{0.49\textwidth}
        \centering
        \includegraphics*[width=\textwidth]{flux-spec-scaled.pdf}
        \caption{}\label{fig:energy-flux}
    \end{subfigure}
% \begin{subcaptiongroup*}
% 	\phantomsubcaption\label{fig:minimum-distance}
% 	\phantomsubcaption\label{fig:prefactors}
% \end{subcaptiongroup*}

    \caption{\emph{(a):} Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines) for $T=1.9K$ (red)
and $T=2.1K$ (blue).\emph{(b):} Spectral normal fluid kinetic energy flux, $\Pi (k)$. It is normalised using the integral scale and the normal fluid energy at reconnection.
\emph{Inset:} Post reconnection evolution of the integral length scale, $L$.
Times and temperatures are labelled as in Fig.~3a.}
\label{fig:spectra}
\end{figure}
%

It clearly emerges that, during the reconnection, energy is predominantly injected into the normal fluid at intermediate and small length scales. 
For $k>5$ in correspondence of the reconnection time $t_R$,
we observe a significant increase of the normal fluid energy spectral density:
$E(k,\, t\approx t_R)/E(k,\, t\ll t_R) \approx 10^2$. 
%
In the post-reconnection regime, we simultaneously observe a small 
decrease of the spectrum at intermediate 
and small scales ($k > 5$) and an increase at large scales, 
suggesting the existence of a mechanism by 
which energy generated at small length scales is transferred to larger scales. 
To shed light on this mechanism, as customary for turbulent flows, we analyse the spectral energy flux 
\begin{equation}
    \Pi(k)=\int_{|{\bf p}|<k}  \hat{\mathbf{v}}_n^*\cdot \left [\widehat{(\v_n\cdot\nabla)\v_n} \right ]d {\bf p} +c.c. \;\; ,
\end{equation}
where $\widehat{\cdot}$ indicates the Fourier transform.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*{\NOTE}[1]{\textbf{\color{red}[#1]}}
\newcommand*{\CHANGE}[1]{{\color{magenta}#1}}
\newcommand*{\DEL}[1]{{\color{green}#1}}

@@

\begin{figure*}
	\centering
  \includegraphics[width=0.95 \textwidth]{snapshots-2.pdf}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snap-4.pdf}

	% \end{subfigure} \hfill
  %   %
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-1.pdf}
	% \end{subfigure}
	% \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-2.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-3.pdf}
	% \end{subfigure}
  %   \begin{subfigure}[b]{0.24\textwidth}
	% 	\centering
	% 	\includegraphics*[width=\textwidth]{snaps-hel-4.pdf}
	% \end{subfigure} \hfill
	\caption{
\CHANGE{
Three-dimensional rendering of the time evolution of
an initially orthogonal superfluid vortex configuration 
undergoing a reconnection at $t=t_R$. Temperature is $T = 1.9$K and time is made
dimensionless with 
% (from left to right) $(t-t_R)/\tau_R=-7.47\times10^{-2}$, $-3.09\times10^{-3}$, $8.79\times10^{-4}$ and $5.39\times10^{-3}$,
$\tau_R = L^2/\kappa$. }
Green tubes represent the superfluid vortex lines (the tubes’ radii have been 
greatly exaggerated for visual purpose). In the top sequence,
the blue volume rendering represents the scaled normal fluid enstrophy 
$\bom^2/\bom^2_{max}$. Note the Kelvin wave on the superfluid
vortex at $(t-t_R)/\tau_R\approx0.001$. In the bottom sequence, the red/blue volume
rendering at the same times represent scaled positive/negative normal fluid
helicity.	
%Three-dimensional rendering of an orthogonal vortex configuration 
%undergoing a vortex reconnection at $T=1.9K$. The red tubes represent
%the superfluid vortex
%lines (the tubes' radii have been greatly exaggerated for visual purposes), 
%and the blue volume rendering represents the scaled normal fluid enstrophy 
%$\bom^2/\bom^2_{max}$. In the third panel, note the Kelvin waves on the superfluid vortices.
}
\label{fig:visualisation}
\end{figure*}

@@

In this work, we unveil a similar dynamics occurring in superfluid helium
($^4$He) as a result of vortex reconnections. 
%\red{The onset and development of quantum turbulence is fundamentally driven by such reconnections, which facilitates energy transfer and the breakdown of coherent vortex structures.}  
%Reconnections occur continuously in turbulence: they take place when
%two vortex lines collide and recombine, exchanging heads and tails, 
%altering the overall topology of the flow
%\cite{koplik-levine-1993,bewley-etal-2008,rorai-etal-2016,serafini-etal-2017,galantucci-baggaley-parker-barenghi-2019,villoisUniversalNonuniversalAspects2017,villois2020irreversible}. 
We show that the mutual friction force arising from the  
vortex reconnection is chiral, injecting in the normal fluid prevalently 
helicity of a given sign 
%\red{which is not globally conserved across the reconnection event}. 
Thus, as a consequence of vortex reconnections,
we observe an increase of the chiral imbalance of the quantum fluid, producing a transfer of kinetic energy from small to large scales, similarly to the phenomenology observed in 3D helically-decimated classical flows. 
 Unlike classical fluids, such a chiral imbalance arises naturally as physical process in the normal fluid.% because of vortex reconnections. 
%This is contrast to classical fluid dynamics,  where the helical characteristics of the flows producing 
% an inverse energy transfer require a careful synthetic construction 
% \cite{biferaleInverseEnergyCascade2012a,biferale-etal-2013,sahoo-etal-2017,plunianInverseCascadeEnergy2020a}.

@@

To model superfluid helium dynamics, we employ the recently developed FOUCAULT model
\cite{galantucciNewSelfconsistentApproach2020b}.  In this approach, superfluid vortex lines are
parametrized as one-dimensional space curves  $\s(\xi,t)$, $\xi$ and $t$ being arclength and time respectively, exploiting the large separation of length scales between the vortex core radius, the Lagrangian discretisation along the vortex lines $\Delta\xi$, and the average radius of curvature $R_c$ of the vortex lines. The vortex lines evolve according to the following equation of motion:
\begin{equation}
    \dot{\s}(\xi,t) = \v_s + \frac{\rho_n}{\rho}\left[\v_{ns}\cdot\s'\right]\s' + \beta\s'\times\v_{ns} + \beta'\s'\times\left[\s'\times\v_{ns}\right],
\end{equation}
where $\dot{\s} = \partial\s/\partial t$, $\s' = \partial\s/\partial\xi$ is the unit tangent vector, $\v_n$ and $\v_s$ are the normal fluid and superfluid velocities at $\s$, $\v_{ns} = \v_n-\v_s$, and $\beta,\, \beta'$ are temperature and Reynolds number dependent mutual friction coefficients \cite{galantucciNewSelfconsistentApproach2020b}. The calculation of the superfluid velocity $\v_s$ is performed via the computation of the Biot-Savart integral de-singularised with standard techniques (see Supplementary Material \cite{suppMat}). The normal fluid is described classically using the incompressible ($\nabla\cdot\v_n=0$) 
Navier-Stokes equation
\begin{equation}
    \frac{\partial\v_n}{\partial t} + (\v_n\cdot\nabla)\v_n = -\frac{1}{\rho}\nabla p  + \nu_n\nabla^2\v_n + \frac{\mathbf{F}_{ns}}{\rho_n} \; \; , 
\end{equation}
%
where $\rho_n$ and $\rho_s$ are the normal fluid and superfluid densities,
$\rho=\rho_n + \rho_s$,  $p$ is the pressure,  $\nu_n$ is the kinematic 
viscosity of the normal fluid, and the mutual friction force per unit
volume, $\mathbf{F}_{ns}$, is the line integral of the mutual friction 
force per unit length, $\mathbf{f}_{ns}$ \cite{suppMat}:
%
\begin{equation}
\mathbf{F}_{ns}(\x) = 
\oint_{\mathcal{C}}\delta(\x-\s)\mathbf{f}_{ns}(\s)d\xi,     
\end{equation}
$\mathcal{C}$ representing the entire vortex configuration. 
The regularisation of mutual friction is performed using a physically self-consistent scheme \cite{galantucciNewSelfconsistentApproach2020b}. %, gualtieri2015exact, gualtieri2017turbulence}. 
We consider a periodical box of size $L=2\pi$ (so that  wavevectors are integers).

To study the reconnection dynamics, we consider 
\CHANGE{
two distinct initial conditions at temperatures $T=1.9K$ and $T=2.1K$: (a) a pair of initially orthogonal vortices (a second pair of orthogonal vortices, with
corresponding vortices of each pair having opposite circulation, is included in the computation in order to preserve periodicity along the boundaries \cite{suppMat};  
the vortex pairs are separated by the distance $D_{\ell}$, while each vortex within each pair is initially at distance $d_{\ell}$ to the other vortex, such that $d_{\ell}\ll D_{\ell}$ in order to ensure that the dynamics in the vicinity of the reconnection is dominated by local interactions); (b) a set of 49 Hopf-links, 
each consisting in two, perpendicular, linked vortex rings at varying initial distances from each other \cite{suppMat}.
}

%All results reported hereafter \red{related to the orthogonal reconnection} refer only to the reconnection of one pair of vortices. 
%\red{In our findings we also report similar observations using Hopf-links from Ref.~\cite{stasiak2025experimental}, which we show in Fig.\ref{fig:total-helicity}.}
%\red{In our simulations, the reconnections of the two pairs of vortices occur at different times; here we consider only the vortex dynamics up to and shortly after the first reconnection (see Fig.~\ref{fig:total-helicity}). Hence, all results reported and discussed herafter refer to the first reconnection.}.

The evolution of the reconnection of an initially orthogonal single vortex pair is reported in Fig.~\ref{fig:visualisation}. 
The first row shows the reconnecting superfluid vortices (in green) accompanied by normal fluid structures generated by the mutual friction, 
here displayed as enstrophy rendering $\bom(\x)^2=|\nabla\times \v_n|^2$. Such structures are the signature of the violent irreversible energy transfers in vortex reconnections 
\cite{stasiak2025experimental}. The second row of Fig.~\ref{fig:visualisation} shows the rendering of the local helicity density $H(\x)=\v_n\cdot\bom$: we observe a clear local helicity production, with an abrupt change of sign due to the rearrangement of the vortex topology. Remarkably, during the reconnection process there is a sudden net overall
normal fluid helicity production, as shown in Fig.~\ref{fig:total-helicity} (a). 
\CHANGE{
Indeed, the temporal evolution of the total normal fluid helicity 
$\mathcal{H} = \int_{\mathcal{V}}H(\x)dV$ computed over the entire volume $\mathcal{V}$ shows that at reconnection time $t_R$ there is a significant overall 
injection of helicity of a given sign. In particular, in the reconnections we monitor the injected helicity is negative (positive) for the orthogonal (Hopf-link) 
reconnections.
}
%We will come back to this finding later.
\begin{figure}[t]
    \centering

    %  \includegraphics[width=0.49\textwidth]{helicity_integralScale.pdf}
       \begin{subfigure}{0.49\textwidth}
         \centering
         \includegraphics*[width=\textwidth]{helicity_total.pdf}
    %     \caption{}\label{fig:}
     \end{subfigure}
     \begin{subfigure}{0.49\textwidth}
         \centering
         \includegraphics*[width=\textwidth]{energy-spec-scaled.pdf}
    %     \caption{}\label{fig:}

     \end{subfigure}
\caption{\CHANGE{ \emph(a): Temporal evolution of the total normal fluid helicity $\mathcal{H}$, normalised by the quantum of circulation $\kappa$. 
Bold solid lines refer to orthogonal reconnections, slightly faded lines refer to the set of Hopf-links reconnections. Red color refers to $T=1.9K$, blue to $T=2.1K$.
\emph(b): Normal fluid kinetic energy spectrum $E(k)$ before 
reconnection (dashed lines), at reconnection (solid lines) 
and after reconnection (dotted lines), for a pair of initially orthogonal vortices. Red (blue) lins correspond to  $T=1.9K$ ($T=2.1K$).}
%For superfluid helium, it is proper to make the helicity dimensionless in terms of $\kappa^2$.
}
%\red{The 3D rendering shows the asynchronous nature of the reconnections in the initial configuration, with the two reconnections occurring at different times; same colour scheme as in Fig.~\ref{fig:visualisation}.}                    }
\label{fig:total-helicity}
\end{figure}

We now focus on the time evolution of the normal fluid energy spectrum $E(k)$, defined by 
\begin{equation}
    E = \frac{1}{(2\pi)^3}\int_{\mathcal{V}}\frac{1}{2}|\v_n|^2 dV = \int_0^{\infty}E(k)dk
\end{equation}
where $E$ is the total normal fluid energy and $k$ is the magnitude of the three-dimensional wavenumber. 
%The energy at reconnection $E_R$ is given by $E(t_R)$, where $t_R$ is the time at reconnection.
The energy spectrum $E(k)$ for orthogonal reconnections is displayed in Fig.~\ref{fig:total-helicity} (b) (corresponding plots for the
Hopf-links reconnections show a similar behaviour \cite{suppMat}).
It clearly emerges that, during the reconnection, energy is predominantly injected into the normal fluid at intermediate and small length scales. 
For $kL>30$ in correspondence of the reconnection time $t_R$,
we observe a significant increase of the normal fluid energy spectral density:
$E(k,\, t\approx t_R)/E(k,\, t\ll t_R) \approx 10^2$. 
%
In the post-reconnection regime, we simultaneously observe a small 
decrease of the spectrum at intermediate 
and small scales ($kL > 30$) and an increase at large scales, 
suggesting the existence of a mechanism by 
which energy generated at small length scales is transferred to larger scales.

\begin{figure}
    \centering
    %\includegraphics[width=0.49\textwidth]{energySpec-fluxSpec.pdf}

     \begin{subfigure}{0.49\textwidth}
         \centering
         \includegraphics*[width=\textwidth]{flux-spec-scaled.pdf}
    %     \caption{}\label{fig:kinetic-energy}
     \end{subfigure}
     \begin{subfigure}{0.49\textwidth}
         \centering
         \includegraphics*[width=\textwidth]{integral-scale-scaled.pdf}
    %     \caption{}\label{fig:energy-flux}
     \end{subfigure}
%\begin{subcaptiongroup*}
%	\phantomsubcaption\label{fig:energy-spec}
%	\phantomsubcaption\label{fig:flux-spec}
%\end{subcaptiongroup*}

    \caption{\CHANGE{
    \emph{(a)}: Spectral normal fluid kinetic energy flux, $\Pi (k)$ for orthogonal reconnections. %, normalised using the box size $L$ and the quantum of circulation $\kappa$}. 
    Times and temperatures are labelled as in Fig.~\ref{fig:total-helicity} (b).
    \emph{(b)}: Post reconnection temporal evolution of the integral length scale, $L_0(t)$ scaled by the reconnection time integral scale $L_R=L_0(t_R)$. 
    Slightly faded lines refer to the set of Hopf-links reconnections (left axis) and bold solid lines refer to orthogonal reconnections (right axis). 
    Red (blue) lins correspond to  $T=1.9K$ ($T=2.1K$).}
    }
\label{fig:spectra}
\end{figure}
%
